---
title: "4 SEM Basics"
author: "Ryan McLean"
date: "October 18, 2019"
output: html_document
---

#Setup
If you need to reload the datasets and libraries, you can run the following. 
```{r}
library(devtools)
library(tidyverse)
library(Hmisc)
library(data.table)
library(plyr)
library(psych)
library(summarytools)
library(car)
```
```{r}
data(mtcars)
iceCream <- haven::read_dta("IceCream.dta")
exam <- haven::read_spss ("Examinations.sav")
flavor <- read.csv(file="flavor2.csv", header=FALSE, sep = ",")
relationship <- haven::read_dta("Relationship.dta")
```

#SEM Specific libraries
These are the libraries that contain the easiest functions for SEM related functions
```{r}
library(lavaan)
library(lavaanPlot)
```


#Confirmatory Factor Analysis
Earlier, you created a scale. A scale is a simple average of all the items that you believe measure a specific construct. However, in SEM, you have the flexibility of creating a factor which will weight the influence of the item on the construct. That means that if one question measures our construct (e.g., depression) better than the other items, your model will reflect that. 
```{r}
RelSatModel <- 'satisfaction =~ rsat1 + rsat2 + rsat3 + rsat4' # Specify the model
fit <- cfa(RelSatModel, data = relationship, missing = "FIML")
summary(fit, standardized=TRUE,fit.measures=TRUE,rsq=TRUE)
lavaanPlot(model = fit, node_options = list(shape = "box", fontname = "Helvetica"), 
           edge_options = list(color = "grey"), 
           coefs = TRUE, stand=TRUE)
```

#General SEM Model
The syntax for defining a bigger SEM model is similar and will be shown below. While my model is fairly simple, you can include as many lines as you need in each section to specify your model
```{r}
model <- '
#Latent Variables
  satisfaction =~ rsat1 + rsat2 + rsat3 + rsat4

#Regression Paths (dependent variable is on left, exogenous are on the right separated by +)
  education ~ satisfaction + age + gender

#Covariances/Correlation Paths
  rsat1 ~~ rsat2
'
fit <- sem(model, data = relationship, missing = "FIML")
summary(fit, standardized=TRUE,fit.measures=TRUE,rsquare=TRUE)
lavaanPlot(model = fit, node_options = list(shape = "box", fontname = "Helvetica"), 
           edge_options = list(color = "grey"), 
           coefs = TRUE, stand=TRUE)
```
#Including Bootstrapping in your model
When you run a model that includes mediation/indirect effects, it is necessary to bootstrap, the syntax should help you
```{r}
# Simply add the se and bootstrap options into your sem fit command and set a seed for reproduceable estimates
set.seed(10)
fit <- sem(model, data = relationship, missing = "FIML", se = "bootstrap", bootstrap = 100)
```

#Requesting modification indices
While respecification should always be guided by theory, SEM tools provide a way to estimate which changes may significantly improve model fit, they can be requested.
```{r}
mi <- modindices(fit) # Requires that you have previously fitted an SEM model
mi

#You can also filter by a specific type. To request only error correlations
mi[mi$op == "~~",]
```

# Testing Measurement Invariance
```{r}
library(semTools)
RelSatModel <- '
satisfaction =~ rsat1 + rsat2 + rsat3 + rsat4
satisfaction ~ 0*1 # set the mean of the latent variable at zero
' 
measurementInvariance(model = RelSatModel, data = relationship, group = "gender", missing = "fiml", strict = TRUE)
#optional
weak.invariance   <- cfa(model=RelSatModel, data=relationship, group= "gendp1_1", missing = "fiml", group.equal="loadings")
strong.invariance <- cfa(model=RelSatModel, data=relationship, group= "gendp1_1", missing = "fiml", group.equal=c("loadings", "intercepts"))
strict.invariance <- cfa(model=RelSatModel, data=relationship, group= "gendp1_1", missing = "fiml", group.equal=c("loadings", "intercepts", "residuals"))

summary(weak.invariance, fit.measures=TRUE)
*** 
summary(strong.invariance, fit.measures=TRUE)
***
summary(strict.invariance, fit.measures=TRUE)

## Can I identify which item is the problem???
```
#Growth Curves
https://cran.r-project.org/web/packages/growthcurver/vignettes/Growthcurver-vignette.html
#Latent Class Analysis
https://cran.r-project.org/web/packages/sBIC/vignettes/LatentClassAnalysis.pdf 
#Moderation and Mediation
The best website for these processes is available here
https://ademos.people.uic.edu/Chapter14.html or if the website is discontinued, it is printed as a pdf in the folder. 
```{r}

```